---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}VaR of a Stock Portfolio}
   \rfoot{\color{headergrey}Module 7}
   \lfoot{\color{headergrey}MSDS 451}
   \fancyfoot[C]{\rmfamily\color{headergrey}Financial and Risk Analytics}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(Rblpapi, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(copula, quietly = TRUE, warn.conflicts = FALSE)
library(fGarch, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(tseries, quietly = TRUE, warn.conflicts = FALSE)
library(gmodels, quietly = TRUE, warn.conflicts = FALSE)
library(rugarch, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/MSDS-RiskAnalytics/datasets/"

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Problem 1

Suppose the risk measure $\Re$ is $VaR(\alpha)$ for some $\alpha$.

Let $P_1$ and $P_2$ be two portfolios whos returns have a joint normal distribution with means $\mu_1$ and $\mu_2$, standard deviations $\sigma_1$ and $\sigma_2$, and correlation $\rho$.

Suppose the initial investments are $S_1$ and $S_2$.

Show that $\Re(P_1 + P_2) <= \Re(P_1) + \Re(P_2)$ under joint normality.

#### Solution:

From __19.12__, we know that:

\setlength{\leftskip}{2cm}

$\widehat{VaR}_{P_i}^{norm}(\alpha) = -S_i \ * \{\hat\mu_{P_i} + \phi^{-1}(\alpha)\hat\sigma_{P_i}\}$

\setlength{\leftskip}{0pt}

Where $S_i$ is the initial value of the portfolio $P_i$ for $i$ in 1, 2. Let the portfolio weights, w,

\setlength{\leftskip}{2cm}
   
$w_1 = \frac{S_1}{S_1 + S_2}$

$w_2 = \frac{S_2}{S_1 + S_2}$

\setlength{\leftskip}{0pt}

Such that,

\setlength{\leftskip}{2cm}

$\mu_P = w_1\mu_1 + w_2\mu_2$

$\sigma^2_p = w_1\sigma^2_1 + w_2\sigma^2_1 + 2w_1w_2\sigma_1\sigma_2\rho$

\setlength{\leftskip}{0pt}

Where $\rho$ < 1, $w_1$ > 0 and $w_2$ > 0.

\setlength{\leftskip}{2cm}

Let $\rho_P < w_1\sigma1 + w_2\sigma_2$

\setlength{\leftskip}{0pt}

When $\alpha < 0.5$, then $\Phi^{-1}(\alpha) < 0$, so that

\setlength{\leftskip}{2cm}

$\Phi^{-1}(\alpha)\sigma_p > w_1\Phi^{-1}(\alpha)\sigma_1 + w_2\Phi^{-1}(\alpha)\sigma_2$

\setlength{\leftskip}{0pt}

Adding portfolio means from above,

\setlength{\leftskip}{2cm}

$\mu_p + \Phi^{-1}(\alpha)\sigma_p > w_1\mu_1 + w_1\Phi^{-1}(\alpha)\sigma_1 + w_2\mu_2 + w_2\Phi^{-1}(\alpha)\sigma_2$

$\ldots = w_1(\mu_1 + \Phi^{-1}(\alpha)\sigma_1) + w_2(\mu_2 + \Phi^{-1}(\alpha)\sigma_2)$

\setlength{\leftskip}{0pt}

Multiplying both sides by -S,

\setlength{\leftskip}{2cm}

$-S\{\mu_p + \Phi^{-1}(\alpha)\sigma_p\} < -Sw_1(\mu_1 + \Phi^{-1}(\alpha)\sigma_1) - Sw_2(\mu_2 + \Phi^{-1}(\alpha)\sigma_2)$

$\ldots = -S_1(\mu_1 + \Phi^{-1}(\alpha)\sigma_1) - S_2(\mu_2 + \Phi^{-1}(\alpha)\sigma_2)$

\setlength{\leftskip}{0pt}

Again using __19.12__,

\setlength{\leftskip}{2cm}

$VaR_{p}^{norm}(\alpha) < VaR_{p_1}^{norm}(\alpha) + VaR_{p_2}^{norm}(\alpha)$

\setlength{\leftskip}{0pt}

\newpage

#### Problem 2

```{r, stock_prices, echo = T, fig.height = 4}

data <- read.csv(paste0(data.dir, "Stock_Bond.csv"),
                 header = T)

prices <- as.matrix(data[1:500, c(3, 5, 7, 9, 11)])

n <- nrow(prices)

returns <- apply(prices, 2, FUN = function( p ){
   (p[-1] / p[1:n-1]) - 1
})

```


+ What are the sample mean vector and sample covariance matrix of the 499 returns on these stocks?

```{r, stock_stats, echo = T}

stock.means <- colMeans(returns)
stock.cov <- cov(returns)

```

```{r, stock_means, echo = F}

mean_df <- as.data.frame(stock.means)
colnames(mean_df) <- c("mu")

pretty_kable(mean_df, "Stock Mean Returns", dig = 5)

```

```{r, stock_cov, echo = F}

pretty_kable(stock.cov, "Covariance Matrix", dig = 5)

```

\newpage

+ How many shares of each stock should one buy to invest $50 million in an equally weighted portfolio?

```{r, stock_shares, echo = T, fig.height = 4}

capital <- 5e7

n.stocks <- ncol(prices)
n.shares <- as.data.frame((capital/n.stocks)/prices[500,])

pretty_kable(data.table(Stock = rownames(n.shares), 
                        Shares = comma(n.shares[,1])), "Equal Weighted Shares")

```

```{r, VaR_setup, echo = T, fig.height = 4}

VaR_norm = function( alpha, mu, sgma, S ) {
   - S * ( mu + qnorm(alpha) * sgma  )
}

stock.weight <- 1/n.stocks
port.weights <- rep( stock.weight, n.stocks )

port.ret <- sum(stock.weight * stock.means)
port.var <- t(port.weights) %*% stock.cov %*% port.weights

```

+ What is the one-day VaR(.01)$^*$ for this equally weighted portfolio? 

```{r, VaR_1day, echo = T, fig.height = 4}

var1d <- VaR_norm(0.1, port.ret, sqrt(port.var), S = 100000)

```

One-day _VaR(.01)_ = `r comma(var1d)`

+ What is the five-day VaR(0.1)$^*$ for this portfolio$^*$?

```{r, VaR_5day, echo = T}

port.ret <- 5 * sum( stock.weight * stock.means )
port.var <- 5 * sum( (1/5)^2 * diag(stock.cov) )

var5d <- VaR_norm(0.1, port.ret, sqrt(port.var), S = 100000 )

```

Five-day _VaR(.01)_ = `r comma(var5d)`

$^*$ VaR estimates assume normality.
