---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today} 
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Huge\bfseries\color{headergrey}Financial and Risk Analytics}
   \rfoot{}
   \fancyfoot[C]{\rmfamily\color{headergrey} Highly Confidential}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
  html_document:
    df_print: paged
---


```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(Rblpapi, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}

theme_set(theme_gdocs())

```

```{r pander_setup, include = FALSE}

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Module 1
### Monte Carlo Simulations

#### Problem Background:

+ Suppose a hedge fund owns __$1,000,000__ of stock and used _$50,000_ of its own capital and _$950,000_ in borrowed money for the purchase.

+ Suppose that if the value of the stock falls below _$950,000_ at the end of any trading day, then the hedge fund will sell all the stock and repay the loan.
+ This will wipe out its $50,000 investment.

The hedge fund is said to be leveraged __20:1__ since its position is _220_ times the amount of its own capital invested.

+ Suppose that the daily log returns on the stock have a mean of __0.05__/year and a standard deviation of __0.23__/year. 

These can be converted to rates per trading day by dividing by 253 and sqrt(253), respectively.

##### Initialization

```{r, global_env, echo = T}

niter <- 1e5 # number of iterations

seed.capital <- 5e4
initial.investment <- log(1e6)

profit.threshold <- log(1.1e6)
loss.threshold <- log(9.5e5)
target.profit <- 1e5

simulate_market <- function(days) {
  # generate random returns for N days
  r <- rnorm(days, mean = 0.05 / 253,
            sd = 0.23 / sqrt(253))

  cumsum(r) # return the final log price after N days.
}


```

\newpage

### Problem 4
######  What is the probability that the value of the stock will be below $950,000 at the close of at least one of the next 45 trading days?

```{r, sim1, echo = T}

set.seed(2009) # reproducible

# setup storage
outcomes <- list(below = rep(0, niter))

# Simulation: Probability dips below $950,000.
for (i in 1:niter) {

  logPrice = initial.investment + simulate_market(45) # simulate 45 trading days.

  minlogP = min(logPrice) # miniumum price over next 45 days

  outcomes$below[i] = as.numeric(minlogP < loss.threshold)
}

```

```{r, prob4, echo = F}

p4 <- round(mean(outcomes$below), 4) * 100

```

_Probability the value of the stock is below $950,000 at least one of next 45 sessions:_ __`r p4`%__

\newpage

Suppose the hedge fund will:

+ sell the stock for a profit of at least __$100,000__ if the value of the stock rises to at least __$1,100,000__ at the end of one of the first _100_ trading days,

+ sell it for a loss if the value falls below __$950,000__ at the end of one of the first _100_ trading days, 

+ or sell it (for "FMV") after _100_ trading days if the closing price has stayed between _$950,000_ and _$1,100,000_.

```{r, sim2, echo = T}

set.seed(2009) # reproducible

outcomes <- list(above = rep(0, niter),
                 below = rep(0, niter),
                 middle = rep(0, niter),
                 pnl = rep(0, niter),
                 open = rep(0, niter))

for (i in 1:niter) {

  logPrice = initial.investment + simulate_market(100) # simulate 100 trading days.

  suppressWarnings({
    # ignore Inf returned if condition not meet.
    profit.day <- min(which(logPrice >= profit.threshold))
    loss.day <- min(which(logPrice <= loss.threshold))
  })

  # What was the exit condition of the position, hince the final price of the stock?
  daysOpen <- ifelse(profit.day == Inf && loss.day == Inf, length(logPrice),
                       min(profit.day, loss.day))

  outcomes$above[i] <- min(profit.day) < min(loss.day)
  outcomes$middle[i] <- profit.day == Inf && loss.day == Inf
  outcomes$below[i] <- min(loss.day) < min(profit.day)

  # p&l = ending value - initial investment
  outcomes$pnl[i] <- exp(logPrice[daysOpen]) - exp(initial.investment)
  outcomes$open[i] <- daysOpen
}

stopifnot(sum(outcomes$above) + sum(outcomes$below) + sum(outcomes$middle) == niter)

```

\newpage

### Problem 5
######  What is the probability that the hedge fund (strategy) will make a profit of at least _$100,000_?

```{r, prob5, echo = T}

# Probability of profit over $100,000.
prob.profit.target <- sum(outcomes$pnl >= target.profit) / length(outcomes$pnl)

p5 <- round(mean(prob.profit.target), 3) * 100

```

_Probability the hedge fund (strategy) returns over $100,000 in profit:_ __`r p5`%__

### Problem 6
######  What is the probability that the hedge fund (strategy) will suffer a loss?

```{r, prob6, echo = T}

# Probability of loss
prob.loss <- sum(outcomes$below) / length(outcomes$below)

p6 <- round(mean(prob.loss), 3) * 100

```

_Probability the hedge fund (strategy) returns a loss:_ __`r p6`%__

### Problem 7
######  What is the expected profit from this trading strategy?

```{r, prob7a, echo = T}

# floating pnl
floating.pnl <- sum(outcomes$pnl[outcomes$middle == 1])

# expected pnl if we assume:
# above close = $100,000, 
# below close = -50,000, 
# and middle = sell at market price.

avg.pnl <- (sum(outcomes$above) * 100000 +
   sum(outcomes$below) * -50000 +
   floating.pnl) / length(outcomes$pnl)

p7a <- round(mean(avg.pnl), 3)

```

_Expected profit/loss assuming stop limits of $100,000 and -50,000:_ __$`r p7a`__


```{r, prob7b, echo = T}

# Using only market prices and actual pnl.
ev.pnl <- mean(outcomes$pnl) # Expected P&L
p7b <- round(ev.pnl, 2)

```

_Expected profit/loss with market orders:_ __$`r p7b`__

### Problem 8
######  What is the expected return?

```{r, prob8, echo = T}

# Expected Return (TW)
strat.returns <- (ifelse(outcomes$pnl < 0,
    -seed.capital, outcomes$pnl) / seed.capital) / outcomes$open
p8 <- round(mean(strat.returns), 5) * 100

```

_Expected (time-weighted) return of the hedge fund (strategy):_ __`r p8`%__